---
title: "ORDINAL LOGISTIC REGRESSION FOR PUBH DATA EQUITY CHALLENGE"
author: "Hajar Ahmed"
date: "2024-10-29"
output: html_document
---

```{r}
#Load Data:
dat <- read.csv(file.choose())

dat[!is.na(dat$Drinking_Level),]


##Install and import packages:

#install.packages("MASS")
#install.packages("brant")
#install.packages("car")
#install.packages("VGAM")

library(MASS)
library(brant)
library(car)
library(VGAM)

```

According to the CDC, the threshold for excessive drinking is different for both men and women. For a woman, 4 or more drinks constitutes excessive drinking, whereas for men that number is 5. For our response variable, we will classify respondants to the SHAPE questionnaire as non-drinker, a moderate drinker, and an excessive drinker. This gives use three levels to use in our ordinal logistic regression.

Explanatory variables we will be studying include...

citation: stats.oarc.ucla

```{r}
#Assess the Factors that are to be utilized for the model
lapply(dat[, c("Categorized_Gender","Categorized_Sexuality", "Education_Status")], table)
```

```{r}
#Create the model:

#Ordinal Logistic Regression model
dat$Drinking_Level <- factor(dat$Drinking_Level, levels = c("1", "2", "3"), ordered = TRUE)
m1 <- polr(Drinking_Level ~ Categorized_Gender + Categorized_Sexuality+ Education_Status, data = dat, Hess = TRUE)


#Generalized Ordinal Logistic Regression model and summary:
m<- vglm(Drinking_Level ~ Categorized_Gender + Categorized_Sexuality +Education_Status, family=cumulative(parallel=FALSE),data=dat)

summary(m)

```

```{r}
# Check multicollinearity assumption of the Ordinal Logistic model:
vif(m1)
```

```{r}
##Check the proportional odds assumption of the Ordinal Logistic model:
brant::brant(m1)

#if any factor has a probability< .05 they fail to pass the assumption
#Therefore we use the generalized Ordinal Logistic model instead.

```

```{r}
#odds coeficcients of the Generalized Ordered Logistic Model
exp(coef(m))

#Example Interpretation of generalized ordered logistic model:
#A Male is 19% less likley than a woman to to go from nondrinking to moderate drinking

```

```{r}
#Calculating a Confidence intreval for each odds Coeficcient:

confint <- confint(m,level=0.95)
exp_confint<-exp(confint)
exp_confint
```

Correlation Plots <https://stackoverflow.com/questions/52554336/plot-the-equivalent-of-correlation-matrix-for-factors-categorical-data-and-mi>

```{r}

#install.packages("ggcorrplot")
library(ggcorrplot)
library(dplyr)
df=c(dat$Categorized_Sexuality)
model.matrix(Drinking_Level~Categorized_Sexuality+Categorized_Gender, data=dat) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag=FALSE, type="lower", lab=TRUE, lab_size=2)

```
